An Automatic Verification Framework for Software Configurations

### 2016-05-02 OSDI/abs.tex
config: new abs

System failures resulting from configuration errors 
have become the major causes affecting the availability and
reliability of today's software systems.
While many misconfiguration handling techniques,
\eg, misconfiguration checking, troubleshooting and repair, 
have been proposed, 
offering automatic verification for configuration files -- like
what we did to regular programs -- is still an open problem.
This is because 1) software configurations are typically written in
poorly structured and untyped representations, and 2) 
writing specifications or constraints for configuration 
verification is non-trivial.
This paper presents, \app, the first automatic verification framework for
software configurations.
For a given configuration file,
\app first analyzes a dataset sample of configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representations.
Then, \app derives rules and constraints by analyzing
the intermediate representations.
The above two steps have formulated a language model.
Finally, the resulting language model can be used
to verify the given configuration file and capable of 
detecting various tricky errors,
including orderings, value correlation, type errors, and entry missing.
%At the heart of \app lies a novel learning engine, 
%which employs a probabilistic type inference strategy and
%defines an intermediate representation for the rule type.
%With this engine, \app learns various type rules (\eg, orderings,
%value relations, type errors, and user defined rules)
%to verify configuration files of interest.
Our \app framework is highly modular, 
does not rely on the system source code, and
can be applied to any new configuration file type with minimal user input.  
%Our tool, named \app, relies on an abstract representation of language 
%rules to allow for this modularity. 
\ennan{One or two sentences here describe the most important
experimental results. For example, we verify real-world misconfiguration 
files and find 95\% errors are detected.}

>>>>

System failures resulting from configuration errors 
have become the major causes affecting the availability and
reliability of today's software systems.
Although many misconfiguration handling techniques,
\eg, misconfiguration checking, troubleshooting and repair, 
have been proposed, 
offering automatic verification for configuration files -- like
what we did to regular programs -- is still an open problem.
This is because software configurations are typically written in
poorly structured and untyped ``languages'', and 
specifying constraints and rules for configuration 
verification is non-trivial in practice.

This paper presents, \app, the first automatic verification framework for
general software configurations.
Given a target configuration file,
\app first analyzes a dataset containing sample configuration files,
thus generating well-structured and probabilistically-typed 
intermediate representations.
Then, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file.
\app framework is highly modular, 
does not rely on the system source code or templates and
can be applied to any new configuration file type with minimal user input.

\app is capable of detecting various tricky errors,
previous methods cannot detect,
including ordering errors, fine-grained value correlation errors, 
entry missing errors, and environment related errors. 
%Our tool, named \app, relies on an abstract representation of language 
%rules to allow for this modularity. 
\ennan{One or two sentences here describe the most important
experimental results. For example, we verify real-world misconfiguration 
files and find 95\% errors are detected.}

### 2016-05-02 OSDI/intro.tex
config: new abs

Configuration errors are one of the most important root causes of
today's software system failures~\cite{xu15systems, yin11anempirical}.
In practice, misconfiguration problems may lead to security
vulnerabilities, application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.  
Although many efforts tackling misconfigurations have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{wang04automatic, attariyan10automating,
su07autobash, whitaker04configuration}, these efforts rely on a manual 
approach to understand and detect the failure symptoms. 
The main reasons for this are:
1) entries in configuration files are untyped assignments, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

Verifying configuration files have been advocated as a reasonable means
to check the correctness of configuration files of interest.

We propose an approach to the verification of  
configuration files which is based on learning rules about the language 
model for configuration files.

Building a practical framework for checking the type-safety of
configuration files, nevertheless, requires addressing several
challenges. First, given the fact that 
different systems' configuration files use various representations, 
the cost of introducing a completely new configuration
language (or representation) to instead of all these existing format looks 
like an impractical goal. This is not only because administrators need 
to learn this new language, but also changes existing system
infrastructure to support such a new language. 
In order to avoid the above issues, we propose a representation,
which is a type-based representation and 
much more structured than existing configuration format, 
but use it as an intermediate layer. We, at the same time,
develop many pluggable parsers
that can transform different systems' configuration representations
into our proposed intermediate representation for the post process.

Second, when running type inference on any given configuration file, 
the type of a variable cannot always be fully determined from a single 
value. For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

@@

Finally, from a systems perspective this is the first approach that {\emph{proactively}} checks 
 the correctness of configuration files. All previous work
~\cite{xu15systems,zhang14encore,yuan11context, wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration} tries to identify the problem after the
failure occured. Our approach isolates potential errors before the system failure occurs, e.g. before the installation. We can also see \app as a tool that can run in conjunction with existing tools. Pre-analyzed configuration files are already free from language-based errors, and this way the workloads of post-failure forensics at the runtime
is significantly reduced, thus making these tools truly practical.

>>>>

Configuration errors are one of the most important root causes of
today's software system failures~\cite{xu15systems, yin11anempirical}.
In practice, misconfiguration problems may lead to security
vulnerabilities, application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.  
For example, Yin {\em et al.}~\cite{yin11anempirical} reveal
that about 31\% system failures are caused by misconfiguration problem.

Although many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{wang04automatic, attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still an open problem.
The main reasons for this are:
1) entries in configuration files are untyped assignments, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

%Offering automatic verification to configuration files -- like
%what we did to programs -- has been advocated as a reasonable means
%to check the correctness of configuration files of interest.
%Nevertheless, it is still an open problem, because 
%1) software configurations are typically written in poorly structured 
%and untyped languages, and 2) writing specifications or constraints
%for configuration verification is non-trivial in practice.

This paper presents \app, the first automatic verification framework
for general software configurations.
The insight of \app design is that we first analyze and learn 
a dataset sample of configuration files, thus building 
a structured and typed language model. Using this ``learned'' language
model, we can verify the target configuration files.
Such a design insight overcomes the above obstacles:
1) learned language model itself is well-structured and typed, so
that we can parse target configuration files to it and verify these
files; 2) because we learn and derive rules and constraints 
from an existing sample dataset, 
we do not need users to write explicit constraints or specifications.

In general, \app's workflow to verifying a given configuration file
could be looked as a three-step methodology.
First, \app analyzes a dataset containing sample configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file and detect potential errors.

Building such an automatic verification framework for
configuration files, nevertheless, still requires addressing several
challenges. 
First, when inferencing each entry's type in a given configuration file, 
the type of a variable cannot always be fully determined from a single 
value. For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type,
like previous work did~\cite{zhang14encore}, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

Second, \ennan{Here, we need to add a challenge description about
why using learning algorithm is interesting.}

\com{First, given the fact that 
different systems' configuration files use various representations, 
the cost of introducing a completely new configuration
language (or representation) to instead of all these existing format looks 
like an impractical goal. This is not only because administrators need 
to learn this new language, but also changes existing system
infrastructure to support such a new language. 
In order to avoid the above issues, we propose a representation,
which is a type-based representation and 
much more structured than existing configuration format, 
but use it as an intermediate layer. We, at the same time,
develop many pluggable parsers
that can transform different systems' configuration representations
into our proposed intermediate representation for the post process.}

@@

\com{
Finally, from a systems perspective this is the first approach that {\emph{proactively}} checks 
 the correctness of configuration files. All previous work
~\cite{xu15systems,zhang14encore,yuan11context, wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration} tries to identify the problem after the
failure occured. Our approach isolates potential errors before the system failure occurs, e.g. before the installation. We can also see \app as a tool that can run in conjunction with existing tools. Pre-analyzed configuration files are already free from language-based errors, and this way the workloads of post-failure forensics at the runtime
is significantly reduced, thus making these tools truly practical.
}
